Short-time simulations of open collective quantum spin systems

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
general revision of intro, theory, and squeezing sections; moved material in section IV to an appendix, wrote/motivated a new section IV on two-time correlators added two-time correlator figure

\title{Structure constants and short-time simulations of open
  collective quantum spin systems}

@@

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum, he2019engineering} that allow for measurement errors
$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
states typically requires nonlinear dynamics that are realized using
collisional, photon-mediated, or phonon-mediated interactions.
Although a truly collective spin model requires uniform, all-to-all
interactions, as long as measurements do not distinguish between
constituent particles, even non-uniform systems can be effectively
described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin operators in the form
\begin{align}
  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
\end{align}
The time evolution of a Heisenberg operator $\O\p{t}$ under a
Hamiltonian $H$ is then given by
\begin{align}
  \f{d}{dt} \O\p{t}
  = T \O\p{t}
  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
\end{align}
for a time derivative (super-)operator $T$ with matrix elements
$T_{\v m\v n}$, determined by
\begin{align}
  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J} \O,
  \label{eq:time_deriv}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$; and is $\D\p{\J}$
is a Heisenberg-picture dissipator, or Lindblad superoperator, defined
by
\begin{align}
  \D\p{\J} \O
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
\end{align}
Decoherence via uncorrelated decay of individual spins at a rate
$\gamma_-$, for example, would be described by the set of jump
operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
can be computed by expanding the product
$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ with
structure constants $f_{\v\ell\v m\v n}$ that are worked out in
Appendices \ref{sec:identities}--\ref{sec:prod_general}, and the
effects of decoherence from jump operators (i.e.~elements of $\J$) of
the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be some of the main technical
contributions of this work, with potential applications beyond the
short-time simulation method presented here.  These ingredients are
sufficient to compute matrix elements $T_{\v m\v n}$ of the time
derivative operator $T$ in Eq.~\eqref{eq:time_deriv} in most cases of
practical interest.

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O\p{t}
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
system of differential equations defined by Eq.~\eqref{eq:EOM}, and
allows us to solve it using standard numerical methods.  Some initial
conditions for this system of differential equations, namely
expectation values of collective spin operators with respect to
spin-polarized (Gaussian) states that are generally simple to prepare
experimentally, are provided in Appendix \ref{sec:initial_conditions}.

The truncation strategy in Eq.~\eqref{eq:weight_truncation} has a few
limitations:
\begin{enumerate*}[label=(\roman*)]
\item simulating a system of differential equations for a large number
  of operators can be time-consuming,
\item the weight measure $w$ may need to be chosen carefully, as the
  optimal measure is generally system-dependent,
\item simulation results can only be trusted up to the time at which
  the initial value of operators $\S_{\v m}$ with weight
  $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
  values of interest.\label{pt:limitation}
\end{enumerate*}
The last limitation in particular unavoidably applies in some form to
any method tracking only a subset of all relevant operators.  We
therefore devise an alternate truncation strategy built around
limitation \ref{pt:limitation}.

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative operator $T^k$ are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\O\p{t}}$ to a fixed order $M$.  Note that the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
of matrix elements $T_{\v m\v n}$ that we may need to explicitly
compute roughly in half.

@@

\section{Spin squeezing, benchmarking, and breakdown}

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to dynamically generate
spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
(TAT), and twist-and-turn (TNT) models described by the collective
spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
\begin{align}
  H_{\t{OAT}} &= \chi S_\z^2, \label{eq:OAT} \\
  H_{\t{TAT}} &= \chi \f13 \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x, \label{eq:TNT}
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally implement
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2015fast}.

@@

The main lesson from Figure \ref{fig:benchmarking} is that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
essentially exact results right up until a sudden and drastic
departure that is simple to diagnose by inspection.  This drastic
departure occurs because the truncated Taylor expansion neglects
high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
contributions to a correlator of interest $\bk{\O\p{t}}$ eventually
become non-negligible.  For sufficiently large times $t$, these
individual contributions may generally be large compared to the value
of $\bk{\O\p{t}}$, in which case their truncation will yield
non-physical results.  This breakdown mechanism is also the reason for
prolonged agreement between the truncated Taylor expansion and
benchmarking methods when squeezing the initial state $\ket{-\Z}$
rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
initial-time correlators $\bk{\S_{\v m}}_{t=0}$ vanish unless
$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
substantially smaller number of neglected terms with non-zero
contribution to the correlator $\bk{\O\p{t}}$.

@@

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
spontaneous decay, excitation, and dephasing of individual spins at
rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that
TNT is much more robust to this kind of decoherence than OAT or TAT, a
finding that could not be deduced from the weak-decoherence
simulations presented in Figure \ref{fig:benchmarking}.
Strong-decoherence computations of this sort were also used to put
lower bounds on theoretically achievable spin squeezing via TAT with
decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the truncated Taylor expansion and the
collective-spin structure constants calculated in this work.

@@

\section{Short-time linear response and two-time commutators}
\label{sec:linear_response}

As another example of physics that is numerically accessible via the
truncated Taylor expansion of Heisenberg operators, we consider the
calculation of two-time commutators, and their relation to the
short-time linear response of Heisenberg operators to perturbations of
a Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by an
operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
operator norm of $\O$, such that the net Hamiltonian is
$\widetilde H=H+V$.  We denote the generator of Heisenberg time
evolution under the perturbed (unperturbed) Hamiltonian by
$\widetilde T$ ($T$).  These generators are related by
\begin{align}
  \widetilde T = T + i\hat V
\end{align}
where $\hat V$ is a superoperator whose action on operators $\O$ is
defined by
\begin{align}
  \hat V \O \equiv \sp{V,\O}_-.
\end{align}
Through quadratic order in the time $t$, we can say that
\begin{align}
  e^{t\widetilde T}
  \approx \f12\sp{e^{it\hat V}, e^{tT}}_+,
\end{align}
which at first order in the perturbation $\hat V$ implies
\begin{align}
  e^{t\widetilde T}
  \approx e^{tT} + \f12 it \sp{\hat V, e^{t T}}_+.
\end{align}
Defining perturbed and unperturbed Heisenberg operators
$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
times $t$ and weak perturbations $V$,
\begin{align}
  \widetilde\O\p{t} - \O\p{t}
  \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
  \label{eq:response}
\end{align}
Short-time linear response functions of this sort,
i.e.~$\widetilde\O\p{t}-\O\p{t}$, can thus be used to experimentally
probe two-time commutators $\sp{V,\O\p{t}}_-$ that contain information
about the spreading of information across many-body degrees of freedom
in a collective spin system.  The numerical methods presented in this
paper can in turn benchmark such experiments, albeit at short times,
even in the presence of strong decoherence.

\newpage

@@

% extension to spin-boson and cavity systems / dicke model / chaos?
% different initial conditions <--> pseudo-steady states in models with a large separation of scales
% OTOCs; theoretical interest; comparing "true" OTOCs against "practical" ones / time reversal of coherent dynamics

Mention/discuss extension of this work to spin-boson systems.

@@

\section{Exact results for the one-axis twisting model}
\label{sec:OAT}

>>>>

\title{Short-time simulations of open collective quantum spin systems}

@@

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
  ma2011quantum}.  Collective spin systems enable a means to break
through this limit via the preparation of many-body entangled states
such as NOON\cite{jones2009magnetic, chen2010heralded,
  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
  lau2014proposal, huang2015quantum}, and spin-squeezed
states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
  appel2009mesoscopic, schleier-smith2010states,
  schleier-smith2010squeezing, leroux2010implementation,
  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
  ma2011quantum, he2019engineering} that allow for measurement errors
$\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$, where
$\varepsilon=1$ saturates the Heisenberg
limit\cite{zwierz2010general}.  Generating such states typically
requires nonlinear dynamics that are realized using collisional,
photon-mediated, or phonon-mediated interactions.  Although a truly
collective spin model requires uniform, all-to-all interactions, as
long as measurements do not distinguish between constituent particles,
even non-uniform systems can be effectively described by a uniform
model with renormalized parameters\cite{hu2015entangled}.

@@

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin operators in the form
\begin{align}
  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
\end{align}
The time evolution of a Heisenberg operator $\O\p{t}$ is then
determined by
\begin{align}
  \f{d}{dt} \O\p{t}
  = T \O\p{t}
  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
\end{align}
for a time derivative (super-)operator $T$ with matrix elements
$T_{\v m\v n}$ defined by
\begin{align}
  T \S_{\v n} \equiv i \sp{H, \S_{\v n}}_-
  + \sum_\J \gamma_\J \D\p{\J} \S_{\v n}
  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
  \label{eq:time_deriv}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $H$ is the collective-spin
Hamiltonian; $\J$ is a set of jump operators with a corresponding
decoherence rate $\gamma_\J$; and is $\D\p{\J}$ is a
Heisenberg-picture dissipator, or Lindblad superoperator, defined by
\begin{align}
  \D\p{\J} \O
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
\end{align}
Decoherence via uncorrelated decay of individual spins at a rate
$\gamma_-$, for example, would be described by the set of jump
operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
can be computed by expanding the product
$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ with
structure constants $f_{\v\ell\v m\v n}$ that are worked out in
Appendices \ref{sec:identities}--\ref{sec:prod_general}, and the
effects of decoherence from jump operators (i.e.~elements of $\J$) of
the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be some of the main technical
contributions of this work, with potential applications beyond the
short-time simulation method presented here.  These ingredients are
sufficient to compute matrix elements $T_{\v m\v n}$ of the time
derivative operator $T$ in Eq.~\eqref{eq:time_deriv} in most cases of
practical interest.

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ to spin operators $\S_{\v m}$ with higher ``weight'',
i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O\p{t}
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
high-weight cutoff $W$.  The truncation in
Eq.~\eqref{eq:weight_truncation} closes the system of differential
equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
using standard numerical methods.  Some initial conditions for this
system of differential equations, namely expectation values of
collective spin operators with respect to spin-polarized (Gaussian)
states that are generally simple to prepare experimentally, are
provided in Appendix \ref{sec:initial_conditions}.

The truncation strategy in Eq.~\eqref{eq:weight_truncation} has a few
limitations:
\begin{enumerate*}[label=(\roman*)]
\item simulating a system of differential equations for a large number
  of operators can be time-consuming,
\item the weight measure $w$ may need to be chosen carefully, as the
  optimal measure is generally system-dependent, and
\item simulation results can only be trusted up to the time at which
  the initial value of operators $\S_{\v m}$ with weight
  $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
  values of interest.\label{pt:limitation}
\end{enumerate*}
The last limitation in particular unavoidably applies in some form to
any method tracking only a subset of all relevant operators.  We
therefore devise an alternate truncation strategy built around
limitation \ref{pt:limitation}.

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative operator $T^k$ are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\O\p{t}}$ to a fixed order $M$.  Note that using the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
qof matrix elements $T_{\v m\v n}$ that we may need to explicitly
compute roughly in half.

@@

\section{Spin squeezing, benchmarking, and breakdown}
\label{sec:squeezing}

To benchmark our method for computing collective spin correlators, we
consider three collective spin models known to generate spin-squeezed
states: the one-axis twisting (OAT), two-axis twisting (TAT), and
twist-and-turn (TNT) models described by the collective spin
Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
\begin{align}
  H_{\t{OAT}} &= \chi S_\z^2, \label{eq:OAT} \\
  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x, \label{eq:TNT}
\end{align}
where we include a factor of $1/3$ in the TAT Hamiltonian because it
naturally appears in realistic proposals to experimentally implement
TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
critical value known to maximize the entanglement generation rate of
TNT in the large-$N$ limit\cite{micheli2003manyparticle,
  sorelli2015fast}.

@@

The main lesson from Figure \ref{fig:benchmarking} is that the
truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
essentially exact results right up until a sudden and drastic
departure that is simple to diagnose by inspection.  This drastic
departure occurs because the truncated Taylor expansion neglects
high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
contributions to a Heisenberg operator $\O\p{t}$ of interest
eventually become non-negligible.  For sufficiently large times $t$,
these individual contributions may generally be large compared to the
value of $\bk{\O\p{t}}$, in which case their truncation will yield
non-physical results.  This breakdown mechanism is also the reason for
prolonged agreement between the truncated Taylor expansion and
benchmarking methods when squeezing the initial state $\ket{-\Z}$
rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
initial-time correlators $\bk{\S_{\v m}}_{t=0}$ vanish unless
$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
substantially smaller number of neglected terms with non-zero
contribution to the correlator $\bk{\O\p{t}}$.

@@

Second, the truncated Taylor expansion enables computing spin
correlators in strong-decoherence regimes entirely inaccessible to
other methods.  As an example, Figure \ref{fig:decoherence_strong}
shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
spontaneous decay, excitation, and dephasing of individual spins at
rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that
TNT is much more robust to this kind of decoherence than OAT or TAT, a
finding that could not be deduced from the weak-decoherence
simulations presented in Figure \ref{fig:benchmarking}.
Strong-decoherence computations of this sort were used to put lower
bounds on theoretically achievable spin squeezing via TAT with
decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
concrete application of the truncated Taylor expansion and the
collective-spin structure constants calculated in this work.

@@

\section{Two-time correlation functions}
\label{sec:two_time}

As a final example of collective-spin physics that is numerically
accessible via the truncated Taylor expansion of Heisenberg operators,
we consider the calculation of collective-spin two-time correlation
functions.  In particular, we consider the effect of decoherence on
the short-time behavior of the two-time connected correlator
\begin{align}
  C\p{t}
  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
  \label{eq:two_time}
\end{align}
in the context of the squeezing models in Section \ref{sec:squeezing}.
In an equilibrium setting, correlation functions of this sort contain
information about the linear response of Heisenberg operators to
perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
setting, they contribute to short-time linear response (see Appendix
\ref{sec:linear_response}).  Similar correlators have made appearances
as order parameters that diagnose time-crystalline phases of
matter\cite{tucker2018shattered}.  Related multi-time correlators,
known as out-of-time-ordered correlators, are commonly examined for
signatures of quantum chaos and information
scrambling[CITE]\cite{CITE}.

\begin{figure*}
  \centering
  \includegraphics{Sp_Sm.pdf}
  \caption{Real (solid lines) and imaginary (dashed lines) parts of
    the two-time connected correlator $C\p{t}$ defined in
    Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
    polarized state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
    evolving under the squeezing Hamiltonians in
    Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.  Results are shown for time
    evolution without decoherence (leftmost panel), as well as with
    decoherence at rates $\gamma_n$ or $\Gamma_n$, respectively
    corresponding to single-spin or collective jump operators
    $s_n^{(j)}$ or $S_n$.  All axes in the sub-figures are shared.  At
    short times, the correlations captured by $C\p{t}$ are suppressed
    by single-spin decay ($\gamma_-$), collective decay ($\Gamma_-$),
    and single-spin dephasing ($\gamma_\z$); these correlations are
    relatively insensitive to collective dephasing ($\Gamma_\z$), and
    can change sign or grow under single-spin ($\gamma_+$) or
    collective ($\Gamma_+$) excitation.}
  \label{fig:two_time}
\end{figure*}

Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
spins, initially in the state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
with and without various sources of decoherence.  The two-time
correlator $C\p{t}$ degrades at short times under certain types of
decoherence, but seems relatively stable to collective dephasing (with
jump operator $S_\z$), and can qualitatively change behavior under in
the presence of single-spin or collective excitation (with jump
operators $s_+^{(j)}$ or $S_+$).  While we do not have an operational
understanding of the correlations captured by $C\p{t}$ in the context
of spin squeezing, Figure \ref{fig:two_time} serves as an example for
the type of quantity and behavior that is accessible at short times
with the truncated Taylor expansion.  This example is straightforward
to extend to multi-time and out-of-time-ordered correlators, as well
as equilibrium settings.

@@

% extension to spin-boson and cavity systems / dicke model / chaos?
% different initial conditions <--> steady states or pseudo-steady states in models with a large separation of scales
% OTOCs; theoretical interest, scrambling; benchmarking "practical" OTOCs (with time-reversal of only coherent dynamics) against "true" OTOCs

@@

\section{Analytical results for the one-axis twisting model}
\label{sec:OAT}
